ProteInfer, deep neural networks for protein functional inference

### 2023-02-28 latex/draft3.tex
errant space

We then assessed the ability of this method to more generally localize function within a sequence, even though the model was not trained with any explicit localization information. We selected all enzymes from Swiss-Prot that had two separate leaf-node EC labels for which our model predicted known EC labels, and these labels were mappable to corresponding Pfam labels. For each of these proteins, we obtained coarse-grained functional localization by using CAM to predict the order of the domains in the sequence and compared to the ground truth Pfam domain ordering\textnormal{(see supplement for details of the method)} %htmlver:(see <a href="./latex/supplement.pdf">supplement</a> for details of the method)
. We found that in 296 of 304 (97\%) of the cases, we correctly predicted the ordering, though we note that the set of bifunctional enzymes for which this analysis is applicable is limited in its functional diversity \textnormal{(see methods)}%htmlver:(see <a href="./latex/supplement.pdf">supplement</a>)
. Although we did not find that fine-grained, per-residue functional localization arose from our application of CAM, we found that it reliably provided coarse-grained annotation of domains' order, as supported by Pfam. This experiment suggests that this is a promising future area for research.
\subsection*{Neural networks learn a general-purpose embedding space for protein function}
Whereas InterProScan compares each sequence against more than 50,000 %from summing integrated signatures at https://www.ebi.ac.uk/interpro/release_notes.html
individual signatures and BLAST compares against an even larger sequence database, ProteInfer uses a single deep model to extract features from sequences that directly predict protein function. One convenient property of this approach is that in the penultimate layer of the network each protein is expressed as a single point in a high-dimensional space. To investigate to what extent this space is useful in examining enzymatic function, we used the ProteInfer EC model trained on the random split to embed each test set protein sequence into a 1100-dimensional vector.

>>>>

We then assessed the ability of this method to more generally localize function within a sequence, even though the model was not trained with any explicit localization information. We selected all enzymes from Swiss-Prot that had two separate leaf-node EC labels for which our model predicted known EC labels, and these labels were mappable to corresponding Pfam labels. For each of these proteins, we obtained coarse-grained functional localization by using CAM to predict the order of the domains in the sequence and compared to the ground truth Pfam domain ordering \textnormal{(see supplement for details of the method)}%htmlver:(see <a href="./latex/supplement.pdf">supplement</a> for details of the method)
. We found that in 296 of 304 (97\%) of the cases, we correctly predicted the ordering, though we note that the set of bifunctional enzymes for which this analysis is applicable is limited in its functional diversity \textnormal{(see methods)}%htmlver:(see <a href="./latex/supplement.pdf">supplement</a>)
. Although we did not find that fine-grained, per-residue functional localization arose from our application of CAM, we found that it reliably provided coarse-grained annotation of domains' order, as supported by Pfam. This experiment suggests that this is a promising future area for research.
\subsection*{Neural networks learn a general-purpose embedding space for protein function}
Whereas InterProScan compares each sequence against more than 50,000 %from summing integrated signatures at https://www.ebi.ac.uk/interpro/release_notes.html
individual signatures and BLAST compares against an even larger sequence database, ProteInfer uses a single deep model to extract features from sequences that directly predict protein function. One convenient property of this approach is that in the penultimate layer of the network each protein is expressed as a single point in a high-dimensional space. To investigate to what extent this space is useful in examining enzymatic function, we used the ProteInfer EC model trained on the random split to embed each test set protein sequence into a 1100-dimensional vector.
